---
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Overview

Rcurvep: an R package for concentration response modeling using Curvep.

The packages provides an R interface for processing concentration response data using Curvep, a response noise filtering algorithm.

With a predefined baseline noise threshold (or minimum response threshold) in an experiment, it allows to calculate activity (with confidence interval) based on original or simulated concentration response data.

If the baseline noise threshold is unknown, the above process can be repeated using a reasonable number of candidates.
The optimal threshold is identified as the the lowest threshold where variance of potency estimation is sufficiently reduced and even stabilized, under the condiction that there are enough response variations induced by chemicals in the dataset.

Currently simulated data can be generated from one of the three types of dataset: 

1. dichotomous binary incidence data (e.g., mortality data from rodent model)
2. continuous data with high number of replicates (e.g., alternative animal model data)
3. continous data with low number of replicates (e.g., in vitro data)

## Installation

```{r, eval = FALSE}
# the development version from GitHub:
# install.packages("devtools")
devtools::install_github("moggces/Rcurvep")
devtools::install_github("moggces/Rcurvep", build_vignettes = TRUE)
```

## Usage

To learn more about Rcurvep, start with the vignettes:
`browseVignettes(package = "Rcurvep")`

