---
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "README_figs/README-")
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Overview

  The package provides an R interface for processing **concentration-response datasets** using Curvep, a response noise filtering algorithm, and other parameteric fitting approaches (e.g., Hill equation). Also, methods for calculating the confidence interval around the activity metrics are also provided. The methods are based on the bootstrap approach to get simulated datasets. The simulated datasets can be used to derive the noise threshold (or benchmark response, BMR) in an assay endpoint. This threshold is critical in the toxicity studies to derive the point-of-departure (POD).


## Installation

```{r, eval = FALSE, echo = TRUE}
# the development version from GitHub:
# install.packages("devtools")
devtools::install_github("moggces/Rcurvep")
devtools::install_github("moggces/Rcurvep", build_vignettes = TRUE)
```

## Usage

### Run analysis

```{r, warning=FALSE, message=FALSE,  echo = TRUE}

library(Rcurvep)
data("zfishbeh")
out_curvep <- combi_run_rcurvep(zfishbeh, TRSH = 30)  # using Curvep with BMR = 30
out_fit <- run_fit(zfishbeh) 

```


### Find BMR

```{r, warning=FALSE, message=FALSE,  echo = TRUE}
data("zfishdev_act")
out_bmr <- estimate_dataset_bmr(zfishdev_act)
```

## More Usage

To learn more about Rcurvep, start with the vignettes:
`browseVignettes(package = "Rcurvep")`

