---
output: 
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Overview

Rcurvep: a R package for concentration response modeling using Curvep.

The packages provides an easy-to-use R interface for processing concentration response data using Curvep, a response noise filtering algorithm.

With a predefined response noise threshold in an experiment, it allows to calculate activity (with confidence interval) based on original or simulated concentration response data.

1. binary incidece data (e.g., mortality data)
2. high number of replicates (e.g., alternative animal model data)
3. low number of replicates (e.g., in vitro data)

If the response noise threshold is unknown, the above process can be repeated using a reasonable number of options.
The optimal threshold is identified as the one that varience of potency estimation is sufficiently reduced and even stabilized.

## Installation

```{r, eval = FALSE}
# the development version from GitHub:
# install.packages("devtools")
devtools::install_github("Rcurvep")
devtools::install_github("Rcurvep", build_vignettes = TRUE)
```

## Usage

To learn more about Rcurvep, start with the vignettes:
`browseVignettes(package = "Rcurvep")`

