---
title: "Investigate the minimum number of chemicals in biobide 32 "
output: html_notebook
author: "Jui-Hua Hsieh"
---

## Introduction

When developing an assay for chemical concentration response screening, it is important and useful to understand the noise level of the developing assay. Inspecting the response variation in vehicle control data is a common use approach such as SD, MAD. It is often useful to identify a threshold, above which, the chemical (based on concentration response data) could be considered as active. The SD, MAD metrics are based on a distribution not based on concentration response data and often, it is hard to decide whether x SD or MAD are needed. Also, how to define the outliers in the distribution can greatly shift the SD value. 

For toxicologists, this threshold is particularly important because it could represent the point of departure (potency). Previously, we provided an approach to identify this threshold based on boostrap samples at the concentrations and Curvep (a noise filtering program for concentration response data). The identified threshold is the threshold at which the variance in potency estimation is substantially reduced. 

The method shows its usefulness in working with the NTP 80 compounds screened on the in vitro and alternative animal assays. 
However, questions remain such as if it is needed to do bootstrap (n = 1000) on 80 compounds since it takes time. 
Parellelization could be a solution but it will be also useful to understand if the number of bootstrap sample/compounds can be reduced. 
Reducing the number of chemicals could be particularly useful for the researchers who are developing assays since usually a small number of compounds are used

In one of the assays in the NTP 80 screen, additional 32 compounds were screened. 
The same threshold was identified using either the 32-compound screen or NTP 80 screen. 

## Goal

By controling the two parameters (n of compounds) and (n of bootstrap samples), investigate the stability of threshold identification methods. (i.e., using raw, non fitted values or using fitted values from exponential fitting (starting values from lm fit or starting values from SSasymp ) 

## Datset

The stored bootstrap sample results (n = 1000) for the 32 compounds

## Steps

1. permutate samples 10, 100, 1000
2. permute number of compounds, 5, 10, 15, 20, 25, 30
3. for each combination (boostrap sample x number of compounds), run 1000 times
4. calculate the metrics 
  a) mean, se of the threshold based on raw values or exponential fit)
  b) fraction of thresholds that need to be checked
  c) fraction of bad exponential fit


## load R packages

```{r, message=FALSE, warning=FALSE}
library(here)
library(Rcurvep)
library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
```


## prepare the dataset

```{r}
zfishdev_boot <- readRDS(here("data-raw", "big_dataset", "biobide32_percent_boot_1000.rds"))
bootd <- zfishdev_boot[["percent_affected_96"]]
#make old format as the new one
bootd <- bootd %>%
  rename(threshold = thres) %>%
  separate(dduid, c("endpoint", "chemical", "direction"), sep = "#") %>%
  separate(direction, c("directionality", "direction"), sep = "@") %>%
  select(-directionality)
acts <- extract_curvep_data(bootd, "act") 
```

# Active (23)/ Inactive (9)

```{r}
sum_act <- extract_curvep_data(bootd, "summary", "INVERSE")
sum_act_hit <- sum_act %>% 
  group_by(chemical, endpoint, direction) %>% 
  summarize(hit_confidence = mean(hit_confidence))

actived <- sum_act_hit %>% filter(hit_confidence > 0.5)
inactived <- sum_act_hit %>% filter(hit_confidence <= 0.5)

p1 <- ggplot(sum_act_hit, aes(x = hit_confidence)) + geom_histogram()
p1
```

In total, regardless of the curvep threshold used, there are  `r nrow(actived)` active compounds out of `r nrow(sum_act_hit)` compounds. The inactive compound names are: `r inactived %>% pull(chemical)`.


## Function to subset the stored data based on number of chemicals and number of bootstrap samples

```{r}
subset_boot_data <- function(dd, n_chemical, n_sample) {
  sel_sample_ids <- seq(1:1000)
  if (n_sample != 1000) sel_sample_ids <- sample(1:1000, n_sample)
  
  
  chem_names <- dd %>% pull(chemical) %>% unique()
  sel_chem_names <- sample(chem_names, n_chemical)

  dd_p <- dd %>% 
    filter(chemical %in% sel_chem_names) %>% 
    filter(repeat_id %in% sel_sample_ids)
  acts <- dd_p %>% #extract_curvep_data(dd_p, "act") %>% 
    mutate(POD = ifelse(is.na(POD), conc_highest, POD)) 
  
  result <- acts %>% identify_basenoise_threshold(., plot = FALSE)
  
  return(list(thres_d = result[[1]], thres_o = result[[2]], sel_chem = sel_chem_names, sel_sample = sel_sample_ids))
  
}
```

## save results

```{r}
save_subset_result <- function(n_chemical, n_sample, times, original_data, dir) {
  result <- map(1:times, function(z) {
    result <- subset_boot_data(dd = original_data, n_chemical = n_chemical, n_sample = n_sample)
    })
  saveRDS(result, file.path(dir, paste0(n_chemical, "-", n_sample, "-", times, ".rds")))
}
```

## update the stored result (for v0.5.2)

```{r}
update_perm_rds <- function(filn, indir, outdir) {
   d_l <- readRDS(file.path(indir, filn))
   d_l_c <- d_l
   map(1:length(d_l), function(y) {
     new_comment <- d_l[[y]][[1]] %>% 
       nest(-endpoint, -direction) %>% 
       mutate(knee_out = purrr::map(data, function(x) cal_knee_point(x, "threshold", "pooled_variance")[[2]])) %>% 
       select(-data) %>% 
       unnest()
     d_l_c[[y]][[2]] <- new_comment
    }
   )
   saveRDS(d_l_c, file.path(outdir, paste0(filn, ".rds")))
}
```


## load the stored results 

```{r}
read_perm_rds <- function(filn, dir) {
  
  pars <- str_replace(filn, ".rds", "") %>% 
    str_split(., "-") %>% unlist() %>% as.numeric() %>% 
    set_names(c("n_chemical", "n_perm", "n_times"))
  
  result <- readRDS(file.path(dir, filn)) %>%
    map_df(., ~ .x[[2]], .id = "perm_id") %>%
    mutate(
      n_chemical = pars[["n_chemical"]],
      n_perm = pars[['n_perm']]
    )
  return(result)
}
```

## calculate the metrics

```{r}

calculate_metrics <- function(dd) {
  permd_sum1 <- dd %>%
    group_by(n_chemical, n_perm, endpoint, direction) %>%
    summarise(
      fraction_check = sum(thresComment == "check")/n(),
      fraction_badfit = sum(cor_exp_fit < 0.95 | is.na(cor_exp_fit))/n(),
      fraction_caution = sum(thresComment == "cautionary")/n()
    ) 
  permd_sum2 <- dd %>%
    filter(thresComment != "check") %>%
    group_by(n_chemical, n_perm, endpoint, direction) %>%
    summarize(
       mean_thresDist_raw = mean(thresDist_raw, na.rm = TRUE),
      se_thresDist_raw = sd(thresDist_raw, na.rm = TRUE)/sqrt(n()),
      mean_thresDist_exp = mean(thresDist_exp, na.rm = TRUE),
      se_thresDist_exp = sd(thresDist_exp, na.rm = TRUE)/sqrt(n())
    )
  
  permd_sum <- inner_join(permd_sum1, permd_sum2)
  
  return(permd_sum)
}

```


## Run the analysis

```{r, message=FALSE, warning=FALSE, eval=FALSE}
set.seed(19800925)
paras <- expand.grid(seq(5, 30, by = 5), c(10, 100, 1000)) %>% set_names(c("n_chemical", "n_sample"))

system.time(pmap(paras, save_subset_result, times = 1000, dir = here("data-raw", "big_dataset", "perm_percent"), original_data = acts))

```

## update perm rds files

```{r}
permd1 <- list.files(here("data-raw", "big_dataset", "perm_percent", "v0.5.1"), pattern = "rds")
map_df(permd1, function(x) {update_perm_rds(filn = x, indir = here("data-raw", "big_dataset", "perm_percent", "v0.5.1"), outdir = here("data-raw", "big_dataset", "perm_percent"))})

```



## Calculate the metrics

```{r}
permd1 <- list.files(here("data-raw", "big_dataset", "perm_percent"), pattern = "rds")
permd1 <- map_df(permd1, function(x) {read_perm_rds(filn = x, dir = here("data-raw", "big_dataset", "perm_percent"))})

permd2 <- list.files(here("data-raw", "big_dataset", "perm_percent", "v0.5"), pattern = "rds")
permd2 <- map_df(permd2, function(x) {read_perm_rds(filn = x, dir = here("data-raw", "big_dataset", "perm_percent", "v0.5"))})
```

```{r}

permd_sum1 <- calculate_metrics(permd1) %>% mutate(ver = "SSasymp")
permd_sum2 <- calculate_metrics(permd2) %>% mutate(ver = "lmstart")

permd_sum <- bind_rows(permd_sum1, permd_sum2) %>% group_by(ver, add = TRUE)

```


### plotting: fraction of check vs n_chemical

```{r}
p1 <- ggplot(permd_sum, aes(x = n_chemical, y = fraction_check, color = as.factor(n_perm)))
p1 <- p1 + geom_point() + 
  geom_line() + facet_grid(ver ~ ., scales = "free")
p1
```

### fraction of bad exponential fit vs n_chemical

```{r}
p2 <- ggplot(permd_sum, aes(x = n_chemical, y = fraction_caution, color = as.factor(n_perm)))
p2 <- p2 + geom_point() + 
  geom_line()  +  facet_grid(ver ~ ., scales = "free")
p2
```

## collect mean and se of the thresholds

```{r}
permd_sum_thres <- permd_sum %>%
  select(mean_thresDist_raw, mean_thresDist_exp) %>%
  gather(thres_type, mean_thres, mean_thresDist_raw, mean_thresDist_exp) %>%
  mutate(
    thres_type = str_replace(thres_type, "mean_", "")
  ) %>%
  inner_join(
    permd_sum %>%
      select(se_thresDist_raw, se_thresDist_exp) %>%
      gather(thres_type, se_thres, se_thresDist_raw, se_thresDist_exp) %>%
      mutate(
        thres_type = str_replace(thres_type, "se_", "")
      )
  )
```

```{r}
p3 <- ggplot(permd_sum_thres, aes(x = n_chemical, y = mean_thres, color = as.factor(n_perm)))
p3 <- p3 + geom_point() + 
  geom_line() +
  facet_grid(thres_type ~ ver, scales = "free") + 
  geom_hline(yintercept = 25) + 
  scale_y_continuous(breaks = seq(18, 36, by = 2)) + 
  scale_x_continuous(breaks = seq(5, 80, by = 5)) +
  geom_errorbar(aes(ymin = mean_thres - se_thres, ymax = mean_thres + se_thres), width = 0.2) 
p3
```

## check the fraction of badfit and fraction of check when using bootstrap = 100 

```{r}
permd_sum %>%
  filter(n_perm == 100) %>%
  arrange(fraction_check)
```


```{r}
permd_sum %>%
  filter(n_perm == 100) %>%
  arrange(fraction_badfit)
```


## Summary

1. Using bootstrap sample = 100 produces simlar results as bootstrap sample = 1000
2. if using  bootstrap sample = 100 & number of compounds = 25 or 30, the fraction of check and fraction of badfit could be lower than 5%
3. the threshold from the fit using either lmstart or SSasymp produces different trends ==> not good

